Meeting the Challenges of Modeling Astrophysical Thermonuclear Explosions: \castro, \maestro, and the \amrex\ Astrophysics Suite

### 2017-10-28 astronum_2017/astronum-amrex.tex
fix formatting

X-ray bursts --- the burning of accreted H/He on the surface of a
neutron star --- can be important probes of neutron star structure.
Interpreting observations requires that we understand what we are
seeing, which can be influenced by the products of the burning and how
the burning spreads across the star.  Many efforts have focused on
different aspects of these events.  One-dimensional models capture the
energetics well and inform us about the nucleosynthesis
\cite{woosley-xrb}.  Global models show the importance of rotation in
confining the burning \cite{SPIT_ETAL02}, while models inspired from
atmospheric science can explore the vertical structure
\cite{cavecchi:2012}.  However, the resolution differences from the
scale of the burning to a reasonable fraction of the neutron star
surface has prevented detailed explorations of the burning and how it
feeds back on the flame structure and propagation in resolved
calculations.  Algorithms and computing architectures are starting to
reach the point where we can span the gaps in spatial scales between
these calculations to provide a better understanding of the ignition
and propagation of the burning front, and the nucleosynthesis
produced.

@@

A parallel effort is porting our microphysics --- in particular
reaction networks --- to GPUs.  Our strategy is to do the entire ODE
integration on the GPU, i.e., the data for a patch of zones is passed
to the GPU, all righthand size and Jacobian evaluations and the
timestepping itself is done on the GPU, and once the burning in all
zones is completed, we access the data as needed on the CPU.  To
enable this, we ported our workhorse ODE integrator (VODE~\cite{vode})
to CUDA Fortran. This required extensive rewrites of the internals of
VODE, which was originally written using Fortran 77 syntax unsupported
by CUDA Fortran. Accelerating VODE with CUDA Fortran has proven
successful, and we now see significant performance gains with the CUDA
version of our reaction networks, even for moderate-sized networks (a
13-isotope standard network).  Figure~\ref{fig:cudaode} shows the
speed-ups on a GPU vs.\ single CPU core.

>>>>

X-ray bursts---the burning of accreted H/He on the surface of a
neutron star---can be important probes of neutron star structure.
Interpreting observations requires that we understand what we are
seeing, which can be influenced by the products of the burning and how
the burning spreads across the star.  Many efforts have focused on
different aspects of these events.  One-dimensional models capture the
energetics well and inform us about the nucleosynthesis
\cite{woosley-xrb}.  Global models show the importance of rotation in
confining the burning \cite{SPIT_ETAL02}, while models inspired from
atmospheric science can explore the vertical structure
\cite{cavecchi:2012}.  However, the resolution differences from the
scale of the burning to a reasonable fraction of the neutron star
surface has prevented detailed explorations of the burning and how it
feeds back on the flame structure and propagation in resolved
calculations.  Algorithms and computing architectures are starting to
reach the point where we can span the gaps in spatial scales between
these calculations to provide a better understanding of the ignition
and propagation of the burning front, and the nucleosynthesis
produced.

@@

A parallel effort is porting our microphysics---in particular
reaction networks---to GPUs.  Our strategy is to do the entire ODE
integration on the GPU, i.e., the data for a patch of zones is passed
to the GPU, all righthand size and Jacobian evaluations and the
timestepping itself is done on the GPU, and once the burning in all
zones is completed, we access the data as needed on the CPU.  To
enable this, we ported our workhorse ODE integrator (VODE~\cite{vode})
to CUDA Fortran. This required extensive rewrites of the internals of
VODE, which was originally written using Fortran 77 syntax unsupported
by CUDA Fortran. Accelerating VODE with CUDA Fortran has proven
successful, and we now see significant performance gains with the CUDA
version of our reaction networks, even for moderate-sized networks (a
13-isotope standard network).  Figure~\ref{fig:cudaode} shows the
speed-ups on a GPU vs.\ single CPU core.
